Bayesian model comparison for rare variant association studies of multiple phenotypes

### 2018-03-16 tex/manuscriptbiorxiv.tex
Manuscript and cover letter updates
I updated the biorxiv version and pushed it to biorxiv. I also removed
the figures from manuscript.tex in preparation for submission. Minor
updates to cover letter.

\documentclass[10pt,letterpaper]{article}
%\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Bayesian model comparison for rare variant association studies of multiple phenotypes} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
\author{Christopher DeBoever$^{1}$, Matthew Aguirre$^{1}$, Yosuke Tanigawa$^{1}$, Chris C. A. Spencer$^{2}$, Timothy Poterba$^{3}$, Mark J. Daly$^{4,5}$, Matti Pirinen$^{5,6,7}$\thanks{matti.pirinen@helsinki.fi}, and Manuel A. Rivas$^{1}$\thanks{mrivas@stanford.edu}}
Christopher DeBoever\textsuperscript{1},
Matthew Aguirre\textsuperscript{1},
Yosuke Tanigawa\textsuperscript{1},
Chris C. A. Spencer\textsuperscript{2},
Timothy Poterba\textsuperscript{3},
Mark J. Daly\textsuperscript{3,4},
Matti Pirinen\textsuperscript{5,6,7*},
Manuel A. Rivas\textsuperscript{1*},
\\
\bigskip
\textbf{1} Department of Biomedical Data Science, Stanford University, Stanford, CA, USA
\\
\textbf{2} Genomics plc, Oxford, UK
\\
\textbf{3} Broad Institute of MIT and Harvard, Cambridge, MA, USA
\\
\textbf{4} Analytic and Translational Genetics Unit, Massachusetts General Hospital, Boston, MA, USA
\\
\textbf{5} Institute for Molecular Medicine Finland (FIMM), University of Helsinki, Helsinki, Finland
\\
\textbf{6} Department of Public Health, University of Helsinki, Helsinki, Finland
\\
\textbf{7} Helsinki Institute for Information Technology HIIT and Department of Mathematics and Statistics, University of Helsinki, Helsinki, Finland
% \textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
% \\
\bigskip

@@

\begin{table}[!ht]
%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
%\caption{
%{\bf $\log_{10}$ Bayes Factors for genes highlighted in Figure~\ref{asthma_bma}.}}
\begin{tabular}{|l+l|l|l|l|l|l|}
\hline
{\bf Gene} & {\bf Joint, protective} & {\bf Joint} & {\bf Eosinophil count} & {\bf FVC} & {\bf FEV$_1$} & {\bf Asthma} \\
\thickhline
\textit{IL33}         &              29.4 &   29.3 &             30.6 &  -2.3 &  -2.2 &    8.1 \\
\hline
\textit{CCR3}         &               3.1 &    3.3 &              7.4 &  -1.4 &  -1.6 &   -0.5 \\
\hline
\textit{RP11-39K24.9} &               0.8 &    1.8 &              4.9 &  -0.1 &  -0.4 &    0.3 \\
\hline
\textit{SCMH1}        &               0.5 &    4.7 &             -1.5 &   7.7 &   3.8 &   -0.7 \\
\hline
\textit{MUSTN1}       &               0.4 &    1.1 &             -1.2 &   2.9 &   2.9 &   -0.6 \\
\hline
\textit{ZFAT}         &               0.3 &    1.3 &             -2.0 &   4.7 &   3.1 &   -0.4 \\
\hline
\textit{ELN}          &               0.2 &    2.5 &             -1.0 &   5.6 &   2.9 &   -0.6 \\
\hline
\textit{C14orf39}     &              -0.7 &   -0.0 &             -1.1 &   3.5 &   2.5 &    0.0 \\
\hline
\textit{TMEM110}      &              -0.9 &    1.1 &             -1.0 &   3.3 &   3.1 &   -0.6 \\
\hline
\textit{IL17RA}       &              -4.4 &    3.1 &              7.9 &  -2.7 &  -2.5 &   -1.1 \\
\hline
\textit{IL18RAP}      &              -9.6 &   -0.9 &             -1.0 &  -1.6 &  -1.7 &    3.5 \\
\hline
\textit{ATP2A3}       &             -11.9 &   -0.8 &             -1.2 &  -2.1 &  -2.2 &    5.8 \\
\hline
\textit{FLG}          &             -20.1 &  -17.2 &             -6.0 &  -7.4 &  -8.5 &    4.0 \\
\hline
\end{tabular}
\label{asthma_table}
%\end{adjustwidth}
\end{table}
\paragraph*{S1 Table.}
\label{asthma_table}
{\bf Highlighted genes from asthma analysis.} $\log_{10}$ Bayes Factors for genes highlighted in Figure~\ref{asthma_bma}.

% %% TABLE: GLAUCOMA GENES OF INTEREST
\begin{table}[!ht]
\begin{adjustwidth}{-1.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
%\caption{
%{\bf $\log_{10}$ Bayes Factors for genes highlighted in Figure~\ref{glaucoma_bma}.}}
\begin{tabular}{|l+p{3cm}|p{1cm}|p{1.6cm}|p{3.8cm}|p{3.4cm}|}
\hline
{\bf Gene} & {\bf Joint, protective} & {\bf Joint} & {\bf Glaucoma} & {\bf Intra-ocular pressure, Goldmann-correlated} & {\bf Corneal resistance factor} \\
\thickhline
\textit{ANGPTL7} &              11.9 &  11.9 &      1.7 &                                       10.1 &                       6.2 \\
\hline
\textit{KLHL22}  &               3.5 &   3.5 &     -0.2 &                                        3.9 &                       2.2 \\
\hline
\textit{WNT10A}  &               2.5 &   2.5 &     -0.3 &                                       -0.5 &                       3.5 \\
\hline
\textit{GLI3}    &               1.5 &   1.5 &     -0.3 &                                       -0.7 &                       3.2 \\
\hline
\textit{RNASE2}  &              -0.5 &  -0.5 &      0.4 &                                       -1.0 &                      -1.1 \\
\hline
\textit{FUK}     &              -1.8 &  -1.8 &      2.5 &                                       -2.1 &                      -2.1 \\
\hline
\end{tabular}
\end{adjustwidth}
\end{table}
\paragraph*{S2 Table.}
\label{glaucoma_table}
{\bf Highlighted genes from glaucoma analysis.} $\log_{10}$ Bayes Factors for genes highlighted in Figure~\ref{glaucoma_bma}.

@@

\section*{Acknowledgments}
This research has been conducted using the UK Biobank Resource under Application Number 24983. We thank all the participants in the UK Biobank study. M.P. is financially supported by the Academy of Finland [288509 and 294050]. M.A.R. and C.D. are supported by Stanford University and a National Institute of Health center for Multi- and Trans-ethnic Mapping of Mendelian and Complex Diseases grant (5U01 HG009080). C.D. is supported by a postdoctoral fellowship from the Stanford Center for Computational, Evolutionary, and Human Genomics and the Stanford ChEM-H Institute. Y .T. is supported by Funai Overseas Scholarship from Funai Foundation for Information Technology and the Stanford University Biomedical Informatics Training Program. The primary and processed data used to generate the analyses presented here are available in the UK Biobank access management system (\url{https://amsportal.ukbiobank.ac.uk/}) for application 24983, ``Generating effective therapeutic hypotheses from genomic and hospital linkage data'' (\url{http://www.ukbiobank.ac.uk/wp-content/uploads/2017/06/24983-Dr-Manuel-Rivas.pdf}), and the results are displayed in the Global Biobank Engine (\url{https://biobankengine.stanford.edu}). We would like to thank the Customer Solutions Team from Paradigm4 who helped us implement efficient databases for queries and application of inference methods to the data. M.A.R. and M.P. are paid consultants in Genomics PLC.

>>>>

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Bayesian model comparison for rare variant association studies of multiple phenotypes} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
\author{Christopher DeBoever$^{1}$, Matthew Aguirre$^{1}$, Yosuke Tanigawa$^{1}$, Chris C. A. Spencer$^{2}$, Timothy Poterba$^{3}$, Mark J. Daly$^{4,5}$, Matti Pirinen$^{5,6,7}$\thanks{matti.pirinen@helsinki.fi}, and Manuel A. Rivas$^{1}$\thanks{mrivas@stanford.edu}}
Christopher DeBoever\textsuperscript{1},
Matthew Aguirre\textsuperscript{1},
Yosuke Tanigawa\textsuperscript{1},
Chris C. A. Spencer\textsuperscript{2},
Timothy Poterba\textsuperscript{3},
Carlos D. Bustamante \textsuperscript{1,4},
Mark J. Daly\textsuperscript{3,5},
Matti Pirinen\textsuperscript{6,7,8*},
Manuel A. Rivas\textsuperscript{1*},
\\
\bigskip
\textbf{1} Department of Biomedical Data Science, Stanford University, Stanford, CA, USA
\\
\textbf{2} Genomics plc, Oxford, UK
\\
\textbf{3} Broad Institute of MIT and Harvard, Cambridge, MA, USA
\\
\textbf{4} Department of Genetics, Stanford University, Stanford, CA, USA
\\
\textbf{5} Analytic and Translational Genetics Unit, Massachusetts General Hospital, Boston, MA, USA
\\
\textbf{6} Institute for Molecular Medicine Finland (FIMM), University of Helsinki, Helsinki, Finland
\\
\textbf{7} Department of Public Health, University of Helsinki, Helsinki, Finland
\\
\textbf{8} Helsinki Institute for Information Technology HIIT and Department of Mathematics and Statistics, University of Helsinki, Helsinki, Finland
% \textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
% \\
\bigskip

@@

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
%\caption{
%{\bf $\log_{10}$ Bayes Factors for genes highlighted in Figure~\ref{asthma_bma}.}}
\begin{tabular}{|l+l|l|l|l|l|l|}
\hline
{\bf Gene} & {\bf Joint, protective} & {\bf Joint} & {\bf Eosinophil count} & {\bf FVC} & {\bf FEV$_1$} & {\bf Asthma} \\
\thickhline
\textit{IL33}         &              29.4 &   29.3 &             30.6 &  -2.3 &  -2.2 &    8.1 \\
\hline
\textit{CCR3}         &               3.1 &    3.3 &              7.4 &  -1.4 &  -1.6 &   -0.5 \\
\hline
\textit{RP11-39K24.9} &               0.8 &    1.8 &              4.9 &  -0.1 &  -0.4 &    0.3 \\
\hline
\textit{SCMH1}        &               0.5 &    4.7 &             -1.5 &   7.7 &   3.8 &   -0.7 \\
\hline
\textit{MUSTN1}       &               0.4 &    1.1 &             -1.2 &   2.9 &   2.9 &   -0.6 \\
\hline
\textit{ZFAT}         &               0.3 &    1.3 &             -2.0 &   4.7 &   3.1 &   -0.4 \\
\hline
\textit{ELN}          &               0.2 &    2.5 &             -1.0 &   5.6 &   2.9 &   -0.6 \\
\hline
\textit{C14orf39}     &              -0.7 &   -0.0 &             -1.1 &   3.5 &   2.5 &    0.0 \\
\hline
\textit{TMEM110}      &              -0.9 &    1.1 &             -1.0 &   3.3 &   3.1 &   -0.6 \\
\hline
\textit{IL17RA}       &              -4.4 &    3.1 &              7.9 &  -2.7 &  -2.5 &   -1.1 \\
\hline
\textit{IL18RAP}      &              -9.6 &   -0.9 &             -1.0 &  -1.6 &  -1.7 &    3.5 \\
\hline
\textit{ATP2A3}       &             -11.9 &   -0.8 &             -1.2 &  -2.1 &  -2.2 &    5.8 \\
\hline
\textit{FLG}          &             -20.1 &  -17.2 &             -6.0 &  -7.4 &  -8.5 &    4.0 \\
\hline
\end{tabular}
\label{asthma_table}
\end{adjustwidth}
\end{table}
\paragraph*{S1 Table.}
\label{asthma_table}
{\bf Highlighted genes from asthma analysis.} $\log_{10}$ Bayes Factors for genes highlighted in Figure~\ref{asthma_bma}.

% %% TABLE: GLAUCOMA GENES OF INTEREST
\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
%\caption{
%{\bf $\log_{10}$ Bayes Factors for genes highlighted in Figure~\ref{glaucoma_bma}.}}
\begin{tabular}{|l+p{3cm}|p{1cm}|p{1.6cm}|p{3.8cm}|p{3.4cm}|}
\hline
{\bf Gene} & {\bf Joint, protective} & {\bf Joint} & {\bf Glaucoma} & {\bf Intra-ocular pressure, Goldmann-correlated} & {\bf Corneal resistance factor} \\
\thickhline
\textit{ANGPTL7} &              11.9 &  11.9 &      1.7 &                                       10.1 &                       6.2 \\
\hline
\textit{KLHL22}  &               3.5 &   3.5 &     -0.2 &                                        3.9 &                       2.2 \\
\hline
\textit{WNT10A}  &               2.5 &   2.5 &     -0.3 &                                       -0.5 &                       3.5 \\
\hline
\textit{GLI3}    &               1.5 &   1.5 &     -0.3 &                                       -0.7 &                       3.2 \\
\hline
\textit{RNASE2}  &              -0.5 &  -0.5 &      0.4 &                                       -1.0 &                      -1.1 \\
\hline
\textit{FUK}     &              -1.8 &  -1.8 &      2.5 &                                       -2.1 &                      -2.1 \\
\hline
\end{tabular}
\end{adjustwidth}
\end{table}
\paragraph*{S2 Table.}
\label{glaucoma_table}
{\bf Highlighted genes from glaucoma analysis.} $\log_{10}$ Bayes Factors for genes highlighted in Figure~\ref{glaucoma_bma}.

@@

\section*{Acknowledgments}
This research has been conducted using the UK Biobank Resource under Application Number 24983. We thank all the participants in the UK Biobank study. M.P. is financially supported by the Academy of Finland [288509 and 294050]. C.D.B. and M.A.R. are supported by the GSP Coordinating Center (U24 HG008956). M.A.R. and C.D. are supported by Stanford University and a National Institute of Health center for Multi- and Trans-ethnic Mapping of Mendelian and Complex Diseases grant (U01 HG009080). M.A.R. is a Faculty Fellow at the Stanford Center for Population Health Sciences. C.D. is supported by a postdoctoral fellowship from the Stanford Center for Computational, Evolutionary, and Human Genomics and the Stanford ChEM-H Institute. Y .T. is supported by Funai Overseas Scholarship from Funai Foundation for Information Technology and the Stanford University Biomedical Informatics Training Program (T32 LM012409). The primary and processed data used to generate the analyses presented here are available in the UK Biobank access management system (\url{https://amsportal.ukbiobank.ac.uk/}) for application 24983, ``Generating effective therapeutic hypotheses from genomic and hospital linkage data'' (\url{http://www.ukbiobank.ac.uk/wp-content/uploads/2017/06/24983-Dr-Manuel-Rivas.pdf}), and the results are displayed in the Global Biobank Engine (\url{https://biobankengine.stanford.edu}). We would like to thank the Customer Solutions Team from Paradigm4 who helped us implement efficient databases for queries and application of inference methods to the data. M.A.R. and M.P. are paid consultants in Genomics PLC. CDB is a member of the scientific advisory boards for Liberty Biosecurity, Personalis, 23andMe Roots into the Future, Ancestry.com, IdentifyGenomics, and Etalon and is a founder of C.D.B. Consulting.
